SU($n$) ferromagnets near the Heisenberg point

### 2020-06-25 collective_sun/sun_perturbation.tex
clarified reason for choosing diagonal +/-1 operators for generating excited states

The remainder of this section is devoted to the construction of a
basis for the $r$-magnon manifolds $\M_r$, and to the expansion of
general spin operators $\O\p{\m w,X}$ in this basis.  We begin with a
classification of excited states in the case of SU(2).  A common
starting point is to first classify states by a definite
(non-negative) spin length $S=\set{N/2,N/2-1,\cdots}$, and a spin
projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization
axis.  In fact, the manifold of states with definite fixed spin length
$S$ is precisely the $\p{N/2-S}$-magnon manifold $\M_{N/2-S}$.  The PS
manifold of Dicke states with $S=N/2$ is then $\M_0$, and more
generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Rather than the spin projection $m_\z$, we then identify states by the
occupation numbers $m=\p{m_\up,m_\dn}\in\A_2\p{N}$, where $m_\mu$
denotes the total number of individual spins in state $\mu$, with
$m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.

@@

Throughout our construction of excitation energy eigenstates of the
form $\O\p{\m w,X}\ket\psi$, where $\ket\psi\in\M_0$ is a PS state, we
have thus far primarily considered the choice of scalar coefficients,
$\m w$.  Here, we discuss the choice of operators, $X$, which allows
us to construct a complete and convenient basis for the set of all
excited states.  Our construction of a basis for the set of excited
states begins with the standard basis $\set{\ket{m}:m\in\A_n\p{N}}$
for the PS manifold $\M_0$, where $m\equiv\p{m_1,m_2,\cdots,m_n}$ and
$m_\mu$ is the (integer) occupation number of the single-spin state
$\mu$.  Written in the ``computational basis'' of $N$ multilevel
spins, these states are
\begin{align}
  \ket{m} =
  \f1{\sqrt{\C\p{m}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
  \Pi \ket{\tilde m},
  &&
  \ket{\tilde m} \equiv \bigotimes_\mu \ket{\mu}^{\otimes m_\mu},
  \label{eq:pops_expansion}
\end{align}
where $\C\p{m}$ is a normalization factor equal to the number of terms
in the sum.  The basic idea in constructing a basis for the set of
excited states is to find a set of operators that give us control over
the signs of the terms in the expansion above, which is sufficient
control to generate {\it any} state that is consistent with the
populations $m$.  Framing the problem of generating excited states as
one of controlling signs in \eqref{eq:pops_expansion} makes it clear
that we can freely restrict ourselves to considering only operators
that are $\pm1$ on the diagonal (and zero elsewhere) in the
computational basis.

The set of all operators that are $\pm1$ on the diagonal (in the
computational basis) is spanned by sums and tensor products of the
single-spin operators
$Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$.  Given a list of
non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
$\sum_\mu z_\mu=r$, we thus define the $r$-spin operator
\begin{align}
  X_z \equiv \bigotimes_\mu Z_\mu^{z_\mu},
\end{align}
and nominally use states of the form
$\ket{m,\m w,z}=\O\p{\m w,X_z}\ket{m}$ as our basis for the $r$-magnon
manifold $\M_r$.  There are three points to address concerning the
states $\ket{m,\m w,z}$:
\begin{enumerate*}
\item depending on the value of $m$, some choices of $z$ result in
  trivial states, $\ket{m,\m w,z}=0$;
  \label{pt:trivial}
\item the set of nontrivial states is generally overcomplete, in that
  any given state can be written as a nontrivial linear combination of
  others, and finally
  \label{pt:dependent}
\item even after identifying a linearly independent set of nontrivial
  states, any given pair of states $\ket{m,\m w,y},\ket{m,\m w,z}$
  with $y\ne z$ generally have a non-zero overlap, which is to say
  that they are not orthogonal.
  \label{pt:orthogonal}
\end{enumerate*}
To address the first two points, \ref{pt:trivial} and
\ref{pt:dependent}, we will place restrictions on the values of $z$
that we allow for any given $m$.  The last point, \ref{pt:orthogonal},
will require us to construct a basis of orthonormal states numerically
for any given values of $m,\m w$.

>>>>

The remainder of this section is devoted to the construction of a
basis for the $r$-magnon manifolds $\M_r$, and to the expansion of
general spin operators $\O\p{\m w,X}$ in this basis.  We begin with a
classification of excited states in the case of SU(2).  A common
starting point is to first classify states by a definite
(non-negative) spin length $S=\set{N/2,N/2-1,\cdots}$, and a spin
projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization
axis.  In fact, the manifold of states with definite fixed spin length
$S$ is precisely the $\p{N/2-S}$-magnon manifold $\M_{N/2-S}$.  The PS
manifold of Dicke states with $S=N/2$ is then $\M_0$, and more
generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Rather than the spin projection $m_\z$, we then identify states by the
occupation numbers $m=\p{m_\up,m_\dn}\in\A_2\p{N}$, where $m_\mu$
denotes the total number of spins in the state $\mu$, with
$m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.

@@

Throughout our construction of excitation energy eigenstates of the
form $\O\p{\m w,X}\ket\psi$, where $\ket\psi\in\M_0$ is a PS state, we
have thus far primarily considered the choice of scalar coefficients,
$\m w$.  Here, we discuss the choice of operators, $X$, which allows
us to construct a complete and convenient basis for the set of all
excited states.  Our construction of a basis for the set of excited
states begins with the standard basis $\set{\ket{m}:m\in\A_n\p{N}}$
for the PS manifold $\M_0$, where $m\equiv\p{m_1,m_2,\cdots,m_n}$ and
$m_\mu$ is the (integer) occupation number of the single-spin state
$\mu$.  Written in the ``computational basis'' of $N$ multilevel
spins, these states are
\begin{align}
  \ket{m} =
  \f1{\sqrt{\C\p{m}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
  \Pi \ket{\tilde m},
  &&
  \ket{\tilde m} \equiv \bigotimes_\mu \ket{\mu}^{\otimes m_\mu},
  \label{eq:pops_expansion}
\end{align}
where $\C\p{m}$ is a normalization factor equal to the number of terms
in the sum.  While $\ket{m}$ is the only PS state with occupation
numbers $m$, the set of {\it all} states with occupation numbers $m$
is spanned by states of the form
\begin{align}
  \ket{m,s} = \f1{\sqrt{\C\p{m}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
  s\p{\Pi} \Pi \ket{\tilde m},
  \label{eq:pops_states}
\end{align}
where $s\p{\Pi}\in\set{\pm1}$ for all permutations $\Pi$.  As a
consequence, when generating excited states of the form
$\O\p{\m w,X}\ket{m}$, we can freely restrict ourselves to operators
$X$ that only change the signs of individual terms (associated with
permutations $\Pi$) in \eqref{eq:pops_expansion}, which suffices to
construct any state $\ket{m,s}$ in \eqref{eq:pops_states}.  We
therefore consider single-operators $X$ that have $\pm1$ on the
diagonal (in the computational basis), and zero everywhere else.

The set of all operators that are $\pm1$ on the diagonal is spanned by
sums and tensor products of the single-spin operators
$Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$.  Given a list of
non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
$\sum_\mu z_\mu=r$, we thus define the $r$-spin operator
\begin{align}
  X_z \equiv \bigotimes_\mu Z_\mu^{z_\mu},
\end{align}
and nominally use states of the form
$\ket{m,\m w,z}=\O\p{\m w,X_z}\ket{m}$ as our basis for the $r$-magnon
manifold $\M_r$.  There are three points to address concerning the
states $\ket{m,\m w,z}$:
\begin{enumerate*}
\item depending on the value of $m$, some choices of $z$ result in
  trivial states, $\ket{m,\m w,z}=0$;
  \label{pt:trivial}
\item the set of nontrivial states is generally overcomplete, in that
  any given state can be written as a nontrivial linear combination of
  others, and finally
  \label{pt:dependent}
\item even after identifying a linearly independent set of nontrivial
  states, any given pair of states $\ket{m,\m w,y},\ket{m,\m w,z}$
  with $y\ne z$ generally have a non-zero overlap, which is to say
  that they are not orthogonal.
  \label{pt:orthogonal}
\end{enumerate*}
To address the first two points, \ref{pt:trivial} and
\ref{pt:dependent}, we will place restrictions on the values of $z$
that we allow for any given $m$.  The last point, \ref{pt:orthogonal},
will require us to construct a basis of orthonormal states numerically
for any given values of $m,\m w$.
